---
title: "Day 2: Maps"
output: html_document
---

```{r setup, include=FALSE}
# This chunk of code sets up how the output should be displayed
# feel free to ignore this!
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Imports and starter code

```{r prelim}
library(dplyr)
library(ggplot2)
# scale bars and north arrows
library(ggspatial)
# maps
library(rnaturalearth)
# map data
library(rnaturalearthdata)
# high-resolution maps
library(rnaturalearthhires)
#' simple features' package
library(sf)
# interactive thematic maps
library(tmap)

# Our malaria dataset
df <- read.csv("../../data/Malaria_Dataset_DataViz.csv")

# Where all the map data is located
map_folder <- "../../data/maps"

# Rivers
rivers <- ne_load(
  scale = 10,
  type = "rivers_lake_centerlines",
  destdir = map_folder,
  returnclass = "sf"
)

# Lakes
lakes <- ne_load(
  scale = 10,
  type = "lakes",
  destdir = map_folder,
  returnclass = "sf"
)

# Oceans
oceans <- ne_load(
  scale = 10,
  type = "ocean",
  destdir = map_folder,
  returnclass = "sf"
)
```

# Exercise 1: plot the world

Use the ```rnaturalearth``` and ```ggplot2``` libraries to get the country outlines on a world map

```{r, ex1, message = FALSE, warning = FALSE}
# Use rnaturalearth to get all countries ("ne_countries")
world_ne <- ne_countries(
  scale = "medium", # map scale
  returnclass = "sf" # return a Simple Features dataframe
)

# We use geom_sf from ggplot to draw simple boundaries and outlines
ggplot(data = world_ne) +
  geom_sf() # geom to visualise sf objects
```

# Exercise 2: Kenya plot a Kenya map?

Plot a high-definition map of Kenya supplemented with information on:

 * Scale
 * Orientation (North/South)
 * Coordinates (longitude/latitude)

```{r, ex2, message = FALSE, warning = FALSE}
# Use rnaturalearth to get the country polygon for Kenya
kenya <- ne_countries(
  country = "kenya", # country name
  type = "countries", # country type
  scale = "large", # requires rnaturalhires
  returnclass = "sf"
)

# Basic Kenya map
kenya_map <- ggplot(data = kenya) +
  geom_sf()

kenya_map_detailed <- kenya_map +
  # Add labels
  labs(
    x = "Longitude",
    y = "Latitude",
    title = "Kenya"
  ) +
  # Add the scale on the bottom left (ggspatial)
  annotation_scale(location = "bl", width_hint = 0.4) +
  # Add a north arrow on the bottom left (ggspatial)
  annotation_north_arrow(
    location = "bl",
    which_north = "true",
    pad_x = unit(0.0, "in"),
    pad_y = unit(0.2, "in"),
    style = north_arrow_fancy_orienteering
  )

kenya_map_detailed
```

# Exercise 3: Kenya add more details?

Augment your Kenya plot by adding:

 * a point for 3 major cities: Nairobi, the capital, as well as Mombasa and Nakuru.
 * lakes, rivers, and potential surrounding oceans

You might have to do a bit of googling to get the coordinates 😀

**Hint:** use ```sf``` and the data in the starter code!

```{r, ex3, message = FALSE, warning = FALSE}
# Kenya data for major cities
kenya_city_data <- data.frame(
  city_name = c("Nairobi", "Mombasa", "Nakuru"),
  lat = c(-1.2921, -4.0435, -0.3031),
  lon = c(36.8219, 39.6682, 36.0800)
)

# Bounding box coordinates for Kenya
kenya_coords <- st_bbox(kenya)
xmin <- kenya_coords[1]
xmax <- kenya_coords[3]
ymin <- kenya_coords[2]
ymax <- kenya_coords[4]

kenya_map_very_detailed <- kenya_map_detailed +
  # Add rivers
  geom_sf(data = rivers, colour = "blue", linewidth = 0.2) +
  # Add lakes
  geom_sf(data = lakes, fill = "lightblue") +
  # Add oceans
  geom_sf(data = oceans, fill = "lightblue") +
  # Restict coordinates for rivers/lakes/oceans
  coord_sf(xlim = c(xmin, xmax), ylim = c(ymin, ymax)) +
  # Add points for cities (scatter plot)
  geom_point(
    data = kenya_city_data,
    mapping = aes(x = lon, y = lat),
    colour = "black"
  ) +
  # Add annotation
  geom_text(
    data = kenya_city_data,
    mapping = aes(x = lon, y = lat, label = city_name),
    nudge_y = 0.25,
    color = "black"
  )

kenya_map_very_detailed
```

# Exercise 4: Kenya add survey data?

Last touch to our map! add the coordinates for surveys in our dataset (```df```) performed in Kenya

```{r, ex4, message = FALSE, warning = FALSE}
kenya_df <- df %>% filter(country == "Kenya")

kenya_map_very_detailed +
  # Scatter plot with x = longitude, y = latitude
  geom_point(
    data = kenya_df,
    mapping = aes(x = longitude, y = latitude),
    colour = "red",
    alpha = 0.2,
    size = 1
  )
```

# Exercise 5: thematic maps with ```tmap```

*Warning!* You might have to delete your current environment before doing this exercise and/or do it on a separate R script.

Let's have a look at interactive thematic maps with ```tmap```!

We provide you with some starter code below. ```World``` is a ```data.frame``` with a special column that contains a geometry for each row, in this case polygons.

```{r, ex5_starter, message = FALSE, warning = FALSE}
# Interactive mode
tmap_mode("view")

# Load World Data
data("World")

# Show the first rows
head(World)
```

Try to plot a few economic indicators for African countries. Can we talk about an "economic burden" for malaria in African countries? Is this enough evidence?

```{r, ex5, message = FALSE, warning = FALSE}
# Select Africa
africa <- World %>% filter(continent == "Africa")

# Rough estimation of average PfPr (parasite rate) in each country
df %>%
  group_by(country) %>%
  summarize(PfPr_mean = mean(PfPr)) %>%
  arrange(desc(PfPr_mean))

# You can plot several columns at once with tm_facets
tm_shape(africa) +
  # Plot life expectancy and GDP per capita
  tm_polygons(
    col = c("life_exp", "gdp_cap_est", "economy"),
    palette = list("Blues", "Blues", "Reds"),
    id = "name"
  ) +
  tm_facets(sync = TRUE, ncol = 3)
```