---
title: "Day 4: Interactive visualisation"
output: html_document
---

```{r setup, include=FALSE}
# This chunk of code sets up how the output should be displayed
# feel free to ignore this!
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

source("../config.R")
```

# Imports and starter code

```{r prelim}
library(ggplot2)
library(plotly)
library(shiny)

df <- read.csv("../../data/Malaria_Dataset_DataViz.csv")

df$mtsi <- df$mosquito_temperature_suitability_index / 1000
```

# Exercise 1: ggplot to plotly

Take your plot from Day 1/Exercise 4 (```fig2```) and make it interactive with ```plotly```!

```{r, ex1, message = FALSE, warning = FALSE}
fig1 <- ggplot(df, aes(y = enhanced_vegetation_index, x = precipitation)) +
  geom_point()

# Linear regression
mod <- lm(enhanced_vegetation_index ~ precipitation, df)

fig2 <- fig1 +
  # classic theme
  theme_classic() +
  # change x-axis, y-axis and title format
  theme(
    axis.text.x = element_text(size = 11L),
    axis.text.y = element_text(size = 11L),
    plot.title = element_text(hjust = 0.5) # centre the title
  ) +
  # Add the regression line
  geom_smooth(method = lm, color = "darkred") +
  # Change x-axis, y-axis and title labels
  labs(
    x = "Precipitation [mm]",
    y = "Vegetation index [A.U.]",
    title = paste("R^2:", format(summary(mod)$r.squared, digits = 2L))
  )

plotly_fig <- ggplotly(fig2)

plotly_fig
```

# Exercise 2: build a Shiny app

*It is recommended to use RStudio and R (not RMarkdown) for this exercise.*

Let's start building an interactive app to compare countries using the 4-panel plot from Day 1/Exercise 5.

The app should incorporate the following elements:
* Input: a country from the Malaria dataset
* Output: an interactive 4-panel plot similar to Day 1/Exercise 5

This exercise will be divided into 3 parts: UI side (part 1), processing the input on the server side (part 2), and finally rendering an interactive a plt (part 3)

Rest assured, you'll find some starter code in a separate script called ```day4_app.R```!

## Part 1: UI side

Our UI should consist of a ```fluidPage``` with:
* A title
* A sidebar layout with two panels:
  * A sidebar panel where a user can select a country from ```unique_countries``` (e.g., like a drop-down menu)
  * A main panel containing a ```plotlyOutput```

<p class="hint">
  Title: ```titlePanel``` <br>
  Side bar layout: ```sidebarLayout``` <br>
  Side bar panel: ```sidebarPanel``` <br>
  Drop-down menu: ```selectInput``` <br>
  main panel: ```mainPanel``` <br>
</p>

```{r, ex2.1, message = FALSE, warning = FALSE, eval = FALSE}
unique_countries <- sort(unique(df$country))

ui <- fluidPage(

  # Give the page a title
  titlePanel("Exploring malaria survey data from African countries"),

  # Generate a row with a sidebar
  sidebarLayout(
    # Define the sidebar with one input
    sidebarPanel(
      # Selection input
      selectInput("country", "Country:",
        choices = unique_countries
      ),
      hr(),
      helpText("Select a country using the drop-down menu.")
    ),
    mainPanel(
      plotlyOutput("plot")
    )
  )
)
```

## Part 2: Server side - processing inputs

Our server should first incorporate the input country by filtering ```df``` such that only the rows from that country are retained.

We are also interested in getting the mean PfPR and its SD (in an object called ```sub_df_pfpr_stats```), grouped by year (like in day1).

To check your code, replace ```plotly_empty``` by:
```{r ex2.2 helper, eval = FALSE, echo = TRUE}
first_year <- # CODE HERE

  first_year_mean_pfpr <- # CODE HERE

  validate(
    need(
      FALSE,
      paste0(
        country(),
        ". First sample obtained in ",
        first_year,
        ". Mean PfPr: ",
        first_year_mean_pfpr
      )
    )
  )
```

When running the app, you should get the following for Angola:

*Angola. First sample obtained in 2005. Mean PfPr: 0.669089607*

<p class="hint">
  Remember to make everything that is susceptible to change (such as the country and country data) as ```reactive```! <br>
</p>

```{r, ex2.2, message = FALSE, warning = FALSE}
server <- function(input, output) {
  country <- reactive({
    input$country
  })

  sub_df <- reactive({
    df |> filter(country == country())
  })

  # Get mean and SD of PfPr grouped by year
  sub_df_pfpr_stats <- reactive({
    sub_df() %>%
      group_by(year) %>%
      summarise(PfPr_mean = mean(PfPr), PfPr_sd = sd(PfPr))
  })

  first_year <- sub_df_pfpr_stats()[1, 1]
  first_year_mean_pfpr <- sub_df_pfpr_stats()[1, 2]

  output$plot <- renderPlotly({
    validate(
      need(
        FALSE,
        paste0(
          country(),
          ". First sample obtained in ",
          first_year,
          ". Mean PfPr: ",
          first_year_mean_pfpr
        )
      )
    )
  })
}
```

## Part 3: Server side - visualization

Our UI is ready, our data processing is ready, now onto the plots!

Replace the validate snippet by a ```subplot``` from the ```plotly``` library, incorporating the 4 panels (over 2 rows) from Day 1/Exercise 5.

<p class="hint">
  Remember that the data is ```reactive``` <br>
  Use ggplotly to make the plots interactive <br>
  Use ```titleX = TRUE``` and ```titleY = TRUE``` for <br>
</p>

```{r, ex2.3, message = FALSE, warning = FALSE}
server <- function(input, output) {
  country <- reactive({
    input$country
  })

  sub_df <- reactive({
    df |> filter(country == country())
  })

  # Get mean and SD of PfPr grouped by year
  sub_df_pfpr_stats <- reactive({
    sub_df() %>%
      group_by(year) %>%
      summarise(PfPr_mean = mean(PfPr), PfPr_sd = sd(PfPr))
  })

  output$plot <- renderPlotly({
    fig1 <- ggplot(sub_df(), aes(x = year, fill = method)) +
      geom_bar() +
      theme_minimal() +
      scale_fill_manual(values = colourblind) +
      labs(
        x = "Year",
        y = "# Surveys"
      )

    fig2 <- ggplot(data = sub_df_pfpr_stats()) +
      geom_point(aes(x = year, y = PfPr_mean)) +
      geom_line(aes(x = year, y = PfPr_mean)) +
      geom_errorbar(
        aes(x = year, ymin = PfPr_mean - PfPr_sd, ymax = PfPr_mean + PfPr_sd),
        width = 0.2
      ) +
      labs(
        x = "Year",
        y = "PfPr (%)"
      ) +
      theme_minimal()

    fig3 <- ggplot(sub_df() %>% filter(rural_urban != ""), aes(x = PfPr, color = rural_urban)) +
      geom_density(aes(y = after_stat(scaled))) +
      theme_minimal() +
      labs(
        y = "Density",
        x = "PfPr (%)",
      )

    fig4 <- ggplot(sub_df(), aes(x = elevation, y = precipitation, colour = mtsi)) +
      geom_point() +
      theme_minimal() +
      scale_colour_distiller(palette = "Purples", direction = 1) +
      labs(
        y = "Precipitation [mm]",
        x = "Elevation [m]",
      )

    m <- list(
      l = 50L,
      r = 50L,
      b = 100L,
      t = 100L,
      pad = 4L
    )

    size <- 800L

    plotly_fig <- subplot(
      ggplotly(fig1, width = size, height = size),
      ggplotly(fig2, width = size, height = size),
      ggplotly(fig3, width = size, height = size),
      ggplotly(fig4, width = size, height = size),
      nrows = 2,
      titleX = TRUE,
      titleY = TRUE,
      margin = 0.1
    )

    plotly_fig |> layout(margin = m)
  })
}
```

