---
title: "Day 0: Refresher"
output: html_document
---

# Potentially useful libraries
```{r, message = FALSE}
library(dplyr)
library(lubridate)
```

# Exercise 1: Data wrangling
For this task, we produced a corrupted version of the classic iris dataset called ```iris_corrupted.csv```.

This dataset describes the characteristics of iris flowers through four features:

* Sepal length
* Sepal width
* Petal length
* Petal width

Obviously, these values have to be strictly positive. However, you will observe that some values are missing/incoherent (completely at random).

Clean the dataset by performing median imputation, i.e., replace the missing values by the feature median.

After replacing the missing values, return the median sepal length of flowers whose petal length is greater than or equal to 5.5 cm.

```{r}
impute_median <- function(path) {
  # x: path to a dataset (data/iris_corrupted.csv)
  # return: float

  # ////////////////////////////////////////////////////////////
  #                      YOUR CODE HERE
  # ////////////////////////////////////////////////////////////
}
```

# Exercise 2: Top-5 occurring words
Return the top 5 occurring words (i.e., characters delimited by spaces) in descending order of the number of occurrences.
The text should first be lower-cased, and thus you should only return lowercase words.

```{r}
top5_words <- function(x) {
  # x: character
  # return: list

  # ////////////////////////////////////////////////////////////
  #                      YOUR CODE HERE
  # ////////////////////////////////////////////////////////////
}
```

# Exercise 3: Workdays until Christmas 2023
For a given day in 2023 with the following format "DD-MM"
Calculate the number of WORKdays left until 25-12-2023

```{r}
workdays_until_xmas <- function(day) {
  # day: character (format: DD-MM)
  # return: int

  # ////////////////////////////////////////////////////////////
  #                      YOUR CODE HERE
  # ////////////////////////////////////////////////////////////
}
```

# Test functions
DO NOT MODIFY THE CODE BELOW!

```{r, echo=FALSE, results="asis"}
test <- function(fun, test_inputs, test_outputs) {
  passed <- 0L

  for (i in seq_along(test_inputs)) {
    input <- test_inputs[i]
    output <- test_outputs[i]
    answer <- fun(input)

    if (identical(answer, output)) {
      cat(sprintf('<span style="color:green;">Passed test %d</span><br>', i))
      passed <- passed + 1L
    } else {
      # cat('<span style="color:red;">
      cat(sprintf('<span style="color:red;">Failed test %d</span><br>', i))
      cat(sprintf("Input: %s<br>", input))
      cat(sprintf("Answer: %s<br>", answer))
      cat(sprintf("Expected: %s<br>", output))
    }
  }
  cat(sprintf("<br><b>Tests passed: %s/%s</b><br><br>", passed, length(test_outputs)))
}

test_impute_median <- function() {
  test_inputs <- c("data/iris_corrupted.csv")

  test_outputs <- c(6.9)

  test(impute_median, test_inputs, test_outputs)
}

test_top5_words <- function() {
  test_inputs <- c(
    "a a a a a b b b b c d d d e e e e e e f g g h",
    "e e e e DDD ddd DdD: ddd ddd aa aA Aa, bb cc cC e e e",
    "  ...  ",
    "  '''  ",
    " /  ",
    " ",
    "To be, or not to be, that is the question: Whether 'tis nobler in the mind to suffer The slings and arrows of outrageous fortune, Or to take arms against a sea of troubles And by opposing end them"
  )

  test_outputs <- list(
    c("e", "a", "b", "d", "g"),
    c("e", "ddd", "aa", "cc", "bb"),
    c(""),
    c(""),
    c(""),
    c(""),
    c("to", "the", "and", "be", "of")
  )

  test(top5_words, test_inputs, test_outputs)
}

test_workdays_until_xmas <- function() {
  test_inputs <- c(
    "17-12",
    "16-11",
    "31-10",
    "19-09",
    "01-08",
    "14-07",
    "21-06",
    "04-05",
    "01-04",
    "14-03",
    "14-02",
    "01-01"
  )

  test_outputs <- c(
    5L,
    27L,
    39L,
    69L,
    104L,
    116L,
    133L,
    167L,
    190L,
    204L,
    224L,
    255L
  )

  test(workdays_until_xmas, test_inputs, test_outputs)
}

cat(sprintf("========== Exercise 1: ========== <br>"))
test_impute_median()

cat(sprintf("========== Exercise 2: ========== <br>"))
test_top5_words()

cat(sprintf("========== Exercise 3: ========== <br>"))
test_workdays_until_xmas()
```