---
title: "Day 1: Data Visualisation Fundamentals"
output: html_document
---

```{r setup, include=FALSE}
# This chunk of code sets up how the output should be displayed
# feel free to ignore this!
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Potentially useful libraries

```{r prelim}
library(dplyr)
library(ggplot2)
library(readxl)

df <- read.csv("../../data/Malaria_Dataset_DataViz.csv")

# Colourblind palette
palette <- c(
  "#0173b2", "#de8f05", "#029e73", "#d55e00",
  "#cc78bc", "#ca9161", "#fbafe4", "#949494", "#ece133",
  "#56b4e9"
)
```

# Exercise 1: Data wrangling

Time to have a first glance at the dataset!
Write a function that:
 1. Only selects the numeric columns that do not end with "id"
 2. Creates a table with the following summary statistics:
  * min, Q1, median, mean, Q3, max
  * the number of NaNs
  * the data type (e.g., ```numeric```)

```{r, exercise1, message = FALSE, warning = FALSE}
describe <- function(data) {
  # data: A data frame
  # return: Data frame with summary statistics for 'data'

  # ////////////////////////////////////////////////////////////
  #                      YOUR CODE HERE
  # ////////////////////////////////////////////////////////////
}

describe(df)
```

# Exercise 2: univariate plots

Something should feel off regarding the column ```mosquito_temperature_suitability```.

Let's plot a histogram to check the distribution. Use both ```hist``` and ```geom_histogram```.

What could have happened? Fix the data from ```mosquito_temperature_suitability``` in a new column called ```mts``` such that the summary looks like this:

```
  Min. 1st Qu.  Median    Mean 3rd Qu.    Max.
  0.00   43.67   59.51   58.59   77.14   94.04
```

```{r, exercise_2, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 8}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 3: customising our univariate plot

Let's customise our last plot (this time using the ```mts``` column)!
Incorporate the following elements:
  - A vertical, red, dashed, line for the median
  - Vertical, black, dashed lines for the first and third quartiles
  - Better data-to-ink-ratio: opt for a white background, dashed grid lines, and remove the top and right borders
  - Increase the font size of the x and y ticks to 11

```{r, exercise_3}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```