---
title: "Day 1: Data Visualisation Fundamentals"
output: html_document
css: styles.css
---

```{r setup, include=FALSE}
# This chunk of code sets up how the output should be displayed
# feel free to ignore this!
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Imports and starter code

```{r prelim}
library(dplyr)
library(gridExtra)
library(ggplot2)
library(readxl)

df <- read.csv("../data/Malaria_Dataset_DataViz.csv")

# Colourblind palette
palette <- c(
  "#0173b2", "#de8f05", "#029e73", "#d55e00",
  "#cc78bc", "#ca9161", "#fbafe4", "#949494", "#ece133",
  "#56b4e9"
)
```

# Exercise 1: Data wrangling

Time to have a first glance at the dataset!

1. Print the number of variables and observations in the dataset

2. Print the first 10 rows of the dataset

3. Write a function called ```describe``` that:
 1. Selects the numeric columns that do not end with "id" in ```df```
 2. Creates a table with the following summary statistics:
  * min, Q1, median, mean, Q3, max
  * the number of NaNs
  * the data type (e.g., ```numeric```)

<p class="hint">
  Filtering: ```select_if``` and/or ```select``` can be used (dplyr) <br>
  Table: There are several possible solutions. The simplest is probably to use ```sapply``` (base) for each statistic.
</p>

```{r, ex1, message = FALSE, warning = FALSE}
describe <- function(data) {
  # data: A data frame
  # return: Data frame with summary statistics for 'data'

  # ////////////////////////////////////////////////////////////
  #                      YOUR CODE HERE
  # ////////////////////////////////////////////////////////////
}

describe(df)
```

# Exercise 2: Univariate plots

Something should feel off regarding the column ```mosquito_temperature_suitability_index```.

Let's plot a histogram to check the distribution. Use both ```hist``` and ```geom_histogram```.

What could have happened? Fix the data from ```mosquito_temperature_suitability_index``` in a new column called ```mtsi``` such that the summary looks like this:

Hint: read the metadata!

```
  Min. 1st Qu.  Median    Mean 3rd Qu.    Max.
  0.00   43.67   59.51   58.59   77.14   94.04
```

<p class="hint">
  Histograms: ```hist``` (base graphics), ```geom_histogram``` (ggplot2)
</p>

```{r, ex2, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 8}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 3: customising our univariate plot

Let's customise our last plot (this time using the ```mtsi``` column)!
Incorporate the following elements:
  - Add a vertical, red, dashed, line for the first three quartiles (Q1, median, Q3)
  - Better data-to-ink-ratio: opt for a white background
  - Add dashed grid lines
  - Increase the font size of the x and y ticks to 11

<p class="hint">
  Vertical lines: ```geom_vline``` (ggplot2) <br>
  Minimal theme: ```theme_minimal``` (ggplot2) <br>
  Grid lines and font size: ```theme``` (ggplot2) <br>
</p>

```{r, ex3}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 4: Bivariate plots

Plot a simple scatter plot showing the following:
* x-axis: precipitation
* y-axis: vegetation index

Assign the plot to the variable ```fig1```

Augment ```fig1``` with:
* A linear regression model fit, coloured in dark red
* A title: "R^2 = xxx" where xxx is the value of R^2 from the linear regression model
* Better labels for the x- and y-axes.

Plot the two figures (fig1 and fig2) side by side. Are the results coherent?

<p class="hint">
Scatter plot: ```geom_points``` (ggplot2) <br>
Trend lines: ```geom_smooth``` (ggplot2)
</p>

```{r, ex4, message = FALSE, warning = FALSE, fig.height = 4, fig.width = 8}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 5: Focus on a single country

Find the country with the most observations in this dataset.

Plot 4 subplots (in a 2 x 2 grid) showing the following:
1. The number of surveys done over the years, grouped by surved method
2. The average *Plasmodium falciparum* parasite rate (PfPr), plotted against the year (mean +/- standard deviation)
3. The distribution of PfPr depending on the area type (rural vs. urban)
4. Combine elevation, precipitation and city accessibility together

<p class="hint">
Bar plot: ```geom_bar``` (ggplot2) <br>
Mean and SD: ```group_by``` and ```summarise``` (dplyr) <br>
Error bars: ```geom_errorbar``` (ggplot2) <br>
Distribution: ```geom_histogram``` or ```geom_density``` (ggplot2) <br>
Q4: a scatter plot should be enough, colouring the points by *mtsi*
</p>

```{r, ex5, message = FALSE, warning = FALSE}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 6: Heatmaps

Compare the average usage of different types of insecticide-treated bed nets for Nigeria and its neighbours (Benin, Chad, Niger, Cameroon).

Plot the resulting table in a heatmap, with the value of each heatmap cell printed inside.

<p class="hint">
Filtering: multiple solutions. Ex: ```filter``` to select the countries, then summarise columns starting with itn after grouping by country
Heatmap: multiple solutions, Ex: ```geom_tile``` (ggplt2) <br>
</p>

```{r, ex6, message = FALSE, warning = FALSE}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```