---
title: "Day 2: Maps"
output: html_document
---

```{r setup, include=FALSE}
# This chunk of code sets up how the output should be displayed
# feel free to ignore this!
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{css, echo=FALSE}
/* This chunk allows to set up hints */
.hint {
  visibility: hidden;
}

.hint::before {
  visibility: visible;
  content: "Hint";
  font-weight: bold
}

.hint:hover {
  visibility: visible;
  font-style: italic;
  font-weight: bold
}

.hint:hover::before {
  display: none;
}
```

# Imports and starter code

```{r prelim}
library(dplyr)
library(ggplot2)
# scale bars and north arrows
library(ggspatial)
# maps
library(rnaturalearth)
# map data
library(rnaturalearthdata)
# high-resolution maps
library(rnaturalearthhires)
#' simple features' package
library(sf)
# interactive thematic maps
library(tmap)

map_folder <- "../data/maps"

# Rivers
rivers <- ne_load(
  scale = 10,
  type = "rivers_lake_centerlines",
  destdir = map_folder,
  returnclass = "sf"
)

# Lakes
lakes <- ne_load(
  scale = 10,
  type = "lakes",
  destdir = map_folder,
  returnclass = "sf"
)

# Oceans
oceans <- ne_load(
  scale = 10,
  type = "ocean",
  destdir = map_folder,
  returnclass = "sf"
)
```

# Exercise 1: plot the world

Use the ```rnaturalearth``` and ```ggplot2``` libraries to get the country outlines on a world map

<p class="hint">
Getting country outlines: ```ne_contries``` (rnaturalearth)
Drawing country outlines: ```geom_sf``` (ggplot2)
</p>

```{r, ex1, message = FALSE, warning = FALSE}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 2: Kenya plot a Kenya map?

Plot a high-definition map of Kenya supplemented with information on:
 * Scale
 * Orientation (North/South)
 * Coordinates (longitude/latitude)

<p class="hint">
Getting Kenya outliners: ```ne_countries``` (rnaturalearth)
Adding scale: ```annotation_scale``` (ggspatial)
Adding North arrow: ```annotation_north_arrow``` (ggspatial)
Adding coordinates: Simply label the axes 😀
</p>

```{r, ex2, message = FALSE, warning = FALSE}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 3: Kenya add more details?

Augment your Kenya plot by adding:
 * a point for 3 major cities: Nairobi, the capital, as well as Mombasa and Nakuru.
 * lakes, rivers, and potential surrounding oceans

You might have to do a bit of googling to get the coordinates 😀

<p class="hint">
Plotting geographical features: ```geom_sf``` (ggplot2)
You will need ```st_bbox``` (sf) to get the bounding coordinates of Kenya, such that you only plot the relevant lakes/rivers/oceans with ```coord_sf``` (ggplot2)
Plotting cities: like a scatter plot
Adding city names: ```geom_text``` (ggplot2)
</p>

```{r, ex3, message = FALSE, warning = FALSE}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```


# Exercise 4: Kenya add survey data?

Last touch to our map! add the coordinates for surveys in our dataset (```df```) performed in Kenya

<p class="hint">
Adding coordinates: like a scatter plot
</p>

```{r, ex4, message = FALSE, warning = FALSE}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 5: thematic maps with ```tmap```

*Warning!* You might have to delete your current environment before doing this exercise and/or do it on a separate R script.

*Warning!* You might have to delete your current environment before doing this exercise and/or do it on a separate R script.

Let's have a look at interactive thematic maps with ```tmap```!

We provide you with some starter code below. ```World``` is a ```data.frame``` with a special column that contains a geometry for each row, in this case polygons.

```{r, ex5_starter, message = FALSE, warning = FALSE}
# Interactive mode
tmap_mode("view")

# Load World Data
data("World")

# Show the first rows
head(World)
```

Try to plot a few economic indicators for African countries. Can we talk about an "economic burden" for malaria in African countries? Is this enough evidence?

<p class="hint">
Selecting African data: ```filter``` (dplyr)
Equivalent to ```ggplot``` (ggplot2): ```tm_shape``` (tmap)
Colouring countries by an indicator: ```tm_polygons(c("indicator"), id = "name")``` (tmap)
Subplots: ```tm_facets``` (tmap)
</p>

```{r, ex5, message = FALSE, warning = FALSE}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```