---
title: "Day 3: Networks"
output: html_document
css: styles.css
---

```{r setup, include=FALSE}
# This chunk of code sets up how the output should be displayed
# feel free to ignore this!
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = TRUE)
```

# Introduction

In this set of exercises, we will go through creating a network in ```igraph``` visualising it
using ```ggplot2```, setting node attributes and plotting some features of the network.

First, we read in the network data - the edge list is in the form of a .txt file, so we
use ```read.table``` instead of ```read.csv```.

This network comes from a school that was used for modelling the spread of COVID-19.
The original publication as well as the data can be found here:
<https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3009790>

After reading the data, we drop the repeated edges from the edge list using distinct
in ```dplyr```. Then make sure the ordering of the nodes from the graph vertices matches
the edge list using match.

Finally, assign the node attributes using the ```$``` on ```V(school_graph)```.

```{r prelim}
# Set working directory here if needed!
library(dagitty)
library(dplyr)
library(ggdag)
library(ggraph)
library(ggplot2)
library(igraph)
library(tidygraph)

# Read in network data
nodes <- read.csv("school_nodelist.csv")
edges <- read.table("school_edgelist_copy.txt")

# Drop repeated edges
edges <- edges[, 1:2] %>% distinct(V1, V2)

# Fix node ordering
school_graph <- graph_from_edgelist(as.matrix(edges))
reorder <- match(nodes$ID, V(school_graph))
nodes <- arrange(nodes, reorder)

# Assign the node attributes using the $ on V(school_graph)
V(school_graph)$role <- nodes$Role
```

# Exercise 1: summary statistics

Now look for some summary statistics about the data - how many individuals are there?
How many roles, and which are the different roles? How many individuals of each role are there?

```{r ex1}
# How many different kinds of roles are there in the network

# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////

# How many people are there in each role?

# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 2: edge case

Currently, it's a bit hard to visualise the full graph, because there are two many nodes.

One thing we could try to do is sample the nodes using the sample function applied
to ```V(graph)``` - try doing this with 50 nodes.

Then, use ```subgraph``` in ```igraph``` to generate a subgraph of the full network with these nodes.

If we have set the node attributes previously for our graph, these attributes will carry over to the subgraph!

Plot the resulting subgraph. Colour the nodes according to the role that each of them plays.

<p class="hint">
use ```as.factor``` to convert the roles to factors, and then assign the colors as node attributes using ```V(graph)$color```
<p>

```{r ex2}
# Randomly subsample the nodes from the full graph - set the seed to get the
# same nodes each time!
set.seed(123)
sample_size <- # Set sample size, e.g. 50
  subgraph_nodes <- sample(V(school_graph), sample_size)

# Create subgraph using these nodes

# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////

# Get the corresponding subgraph node attributes, i.e.
# subgraph_node_attrs <- nodes %>% filter(ID %in% subgraph_nodes) %>% select(ID, Role)

# Set the colours of the roles

# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////

# Choose a layout

# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
# Calculate the degrees of your subgraph

# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////

# Plot!

# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 3: network summary statistics

When we sample the vertices, we are excluding full data about the network.

As we discussed in lectures, one way to visualise information about the whole
graph is to show the degree distribution.

Plot the degree distribution for the whole graph. Do this once for every node,
and then repeat with the distributions for each separate role. Do the degree
distributions look significantly different for the different roles?

```{r}
nodes$degree <- degree(school_graph)


# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

Now show the distributions separated by roles. Compare the histogram with the density plot -
which do you prefer?

```{r}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 4: Remember the subgraph?

Let's also check whether the subgraph is a good representation of the full graph in terms
of degree distribution. Plot the degree distributions for the two graphs in a single plot.

Do they look similar?

```{r ex4}
subgraph_node_degrees <- nodes %>%
  filter(ID %in% subgraph_nodes) %>%
  select(ID, degree)

# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 5 (extension): node to self: there are more network libraries

Extension: Have a go at converting the school subgraph into a ```tidygraph``` object. Use the function
as_tbl_graph to convert an igraph object to a ```tidygraph```, and then use ggraph to plot.

Experiment with different ways of displaying the nodes and edges, as well as the layout and
colors of the vertices. Do you prefer this way of visualising networks compared with ```igraph```?


```{r ex5}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 6: DAG-nificent

Let's look at creating a DAG in ```dagitty```: Create a simple DAG with 5 variables,
called A, B, C, D and E, and make up some associations between the variables.

If you prefer, you could make a DAG from your real research!

If you want to, you can also try playing around with the features in ```ggdag``` - this
requires you to make a ```dagify()``` object instead. See this tutorial for some ideas
of how to start <https://cloud.r-project.org/web/packages/ggdag/vignettes/intro-to-ggdag.html>
and the kinds of functions available.

```{r ex6}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```


