---
title: "Day 3: Dimensionality reduction"
output: html_document
css: styles.css
---

```{r setup, include=FALSE}
# This chunk of code sets up how the output should be displayed
# feel free to ignore this!
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

source("../config.R")
```

```{r prelim}
library(corrplot)
library(gridExtra)
library(dplyr)
library(ggplot2)
library(uwot)

# Our malaria dataset
df <- read.csv("../data/Malaria_Dataset_DataViz.csv")

df$mtsi <- df$mosquito_temperature_suitability_index / 1000

cols_of_interest <- c(
  "itnuse0", "itnuse1", "itnuse2", "itnuse3",
  "act", "irs", "tasseled_cap_wetness", "mtsi",
  "precipitation", "enhanced_vegetation_index", "elevation", "nighttime_lights",
  "accessibility_tocity"
)

countries <- c("Sudan", "Zambia", "Nigeria")

set.seed(42)
```

# Exercise 1: correlation matrix

We have selected for you 13 columns of interest, and 3 different African countries from an economic, geographic, and demographic perspective: Sudan, Zambia, and Nigeria.

Make a correlation matrix using the features of interest then plot it.

<p class="hint">
  Correlation matrix: ```cor``` from the ```stats``` package <br>
  Correlation matrix plot: ```corrplot``` from the ```corrplot``` package
</p>

```{r ex1}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 2: PCA

1. Run a PCA on the subset of interest. Do not forget to scale and center the data!
2. Run a scree plot. How many components are worth retaining?

<p class="hint">
  PCA: ```prcomp``` from the ```stats``` package <br>
  Scree plot: ```screeplot``` from the ```stats``` package
</p>

```{r ex2}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 3: more PCA

Run a biplot to represent the information contained in the first two components.

1. Observe the arrows: what do the principle components capture?
2. Colour the biplot points by country: do the surveys cluster well by country? do the clusters make sense?
3. Colour the biplot points by PfPr:

<p class="hint">
  Biplot: ```biplot``` from the ```stats``` package or ```ggbiplot``` from the ```ggbiplot``` package <br>
</p>

```{r ex3}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 4: UMAP

Let's now turn to UMAP.

Perform 4 UMAP plots with 4 different values for ```n_neighbors```: 2, 20, 200, 2000 (it should take ~ 5 min).

What do you notice? How do these visualisations compare to your PCA plot?


<p class="hint">
  UMAP: ```umap``` (uwot) with ```fast_sgd = TRUE```
</p>

```{r ex4}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```
