---
title: "Day 4: Interactive visualisation"
output: html_document
styles: styles.css
---

```{r setup, include=FALSE}
# This chunk of code sets up how the output should be displayed
# feel free to ignore this!
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

source("../config.R")
```

# Imports and starter code

```{r prelim}
library(dplyr)
library(ggplot2)
library(plotly)
library(shiny)

df <- read.csv("../../data/Malaria_Dataset_DataViz.csv")

df$mtsi <- df$mosquito_temperature_suitability_index / 1000
```

# Exercise 1: ggplot to plotly

Take your plot from Day 1/Exercise 4 (```fig2```) and make it interactive with ```plotly```!

```{r, ex1, message = FALSE, warning = FALSE}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

# Exercise 2: build a Shiny app

*It is recommended to use RStudio and R (not RMarkdown) for this exercise.*

Let's start building an interactive app to compare countries using the 4-panel plot from Day 1/Exercise 5.

The app should incorporate the following elements:
* Input: a country from the Malaria dataset
* Output: an interactive 4-panel plot similar to Day 1/Exercise 5

This exercise will be divided into 3 parts: UI side (part 1), processing the input on the server side (part 2), and finally rendering an interactive a plt (part 3)

Rest assured, you'll find some starter code in a separate script called ```day4_app.R```!

## Part 1: UI side

Our UI should consist of a ```fluidPage``` with:
* A title
* A sidebar layout with two panels:
  * A sidebar panel where a user can select a country from ```unique_countries``` (e.g., like a drop-down menu)
  * A main panel containing a ```plotlyOutput```

<p class="hint">
  Title: ```titlePanel``` <br>
  Side bar layout: ```sidebarLayout``` <br>
  Side bar panel: ```sidebarPanel``` <br>
  Drop-down menu: ```selectInput``` <br>
  main panel: ```mainPanel``` <br>
</p>

```{r, ex2.1, message = FALSE, warning = FALSE}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

## Part 2: Server side - processing inputs

Our server should first incorporate the input country by filtering ```df``` such that only the rows from that country are retained.

We are also interested in getting the mean PfPR and its SD (in an object called ```sub_df_pfpr_stats```), grouped by year (like in day1).

To check your code, replace ```plot_empty``` by:
```{r ex2.2 helper, eval = FALSE, echo = TRUE}
first_year <- # CODE HERE

  first_year_mean_pfpr <- # CODE HERE

  validate(
    need(
      FALSE,
      paste0(
        country(),
        ". First sample obtained in ",
        first_year,
        ". Mean PfPr: ",
        first_year_mean_pfpr
      )
    )
  )
```

When running the app, you should get the following for Angola:

*Angola. First sample obtained in 2005. Mean PfPr: 0.669089607*

<p class="hint">
  Remember to make everything that is susceptible to change (such as the country and country data) as ```reactive```! <br>
</p>

```{r, ex2.2, message = FALSE, warning = FALSE}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```

## Part 3: Server side - visualization

Our UI is ready, our data processing is ready, now onto the plots!

Replace the validate snippet by a ```subplot``` from the ```plotly``` library, incorporating the 4 panels (over 2 rows) from Day 1/Exercise 5.

<p class="hint">
  Remember that the data is ```reactive``` <br>
  Use ggplotly to make the plots interactive <br>
  Use ```titleX = TRUE``` and ```titleY = TRUE``` for <br>
</p>

```{r, ex2.3, message = FALSE, warning = FALSE}
# ////////////////////////////////////////////////////////////
#                      YOUR CODE HERE
# ////////////////////////////////////////////////////////////
```